Based on polling as of 27 October 2024, Harris expects to win on 49.2%, leads Trump by 2% in 2024 US election

### 2024-10-30 paper/paper.qmd

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

harris_data <- harris_data %>% mutate(candidate = "Harris")
trump_data <- trump_data %>% mutate(candidate = "Trump")
combined_data_total <- bind_rows(harris_data, trump_data)

selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
recent_date_cutoff <- as.Date("2024-08-01")  # Set cutoff for recent data

harris_filtered <- harris_data %>%
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)

trump_filtered <- trump_data %>%
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)

# Combine datasets for plotting with a new column for candidate
harris_filtered <- harris_filtered %>% mutate(candidate = "Harris")
trump_filtered <- trump_filtered %>% mutate(candidate = "Trump")
combined_data <- bind_rows(harris_filtered, trump_filtered)

combined_data_total|>filter(national_poll==1)|>ggplot(aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Country",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-27"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  )
```

```{r}
ggplot(combined_data, aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  facet_wrap(~ state) +
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Selected States",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-31"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate"))
```

@@

```{r}
# Load necessary libraries
library(boot)    # For bootstrapping
library(caret)   # For cross-validation

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Diagnostics of model using residual vs fitted plot and norm Q-Q plot\n-Support for Harris"
#| fig-subcap: ["Residual Plot for Unweighted Model", "Q-Q Plot for Unweighted Model", "Residual Plot for Weighted Model", "Q-Q Plot for Weighted Model"]
#| layout-ncol: 2

@@

```
Generally, we use Residual vs Fitted plot and Q-Q Plot diagnostic our model. Residual vs Fitted plot aare Residuals (differences between observed and predicted values) plotted against fitted values. Ideally, these residuals should be randomly scattered around the zero line to indicate that the model does not have systematic errors. The Q-Q plot for the unweighted model shows how the residuals align with a theoretical normal distribution. Ideally, residuals should follow a straight line in this plot if they are normally distributed, which is an assumption of linear regression.

>>>>

## Recent 3 Months Support Trends Prediction

The recent three-month support trends for Kamala Harris and Donald Trump across key battleground states—Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin—show a highly competitive race. Each state’s panel displays support percentages over time from August to November 2024, with Harris shown in blue and Trump in red. The plot uses point size to represent poll weights, accounting for factors such as recency, pollster quality, and sample size, where larger points indicate higher-weighted polls, emphasizing more reliable data.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-support-trends-by-state
#| fig-cap: Support Trends for Kamala Harris and Donald Trump Across Selected States

# Load necessary libraries
library(tidyverse)
library(ggplot2)

@@

selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")

recent_date_cutoff <- as.Date("2024-08-01")  # Set cutoff for recent data

harris_filtered <- harris_data %>%
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)|>mutate(candidate = "Harris")
trump_filtered <- trump_data %>%
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)|>mutate(candidate = "Trump")

combined_data <- bind_rows(harris_filtered, trump_filtered)
# Plot with recent data, specific axis formatting, and smoother trend lines
ggplot(combined_data, aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  facet_wrap(~ state) +
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Selected States",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-31"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate"))
```
@fig-support-trends-by-state illustrates the recent support trends for Kamala Harris and Donald Trump across selected battleground states from August to November 2024. Each subplot represents a different state, with support percentages for each candidate displayed over time. Key features include Harris shown in blue and Trump in red, with smoothed trend lines to capture changes over time. The plot's y-axis spans from approximately 30% to 50% support, while the x-axis highlights monthly intervals from August to November. For example, in Arizona, Trump maintains a slight lead with support fluctuating around 50% in October, compared to Harris around 48%. In Georgia, Trump's support reaches just above 49% in late September, whereas Harris hovers closer to 47%. In Wisconsin, both candidates show tightly aligned trends near 48%, highlighting a close race. Each point's size reflects the poll weight—calculated using recency, pollster quality, and sample size—indicating the relative importance of each poll. This visualization emphasizes recent trends in candidate support across key battleground states.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-national-support-trends
#| fig-cap: Support Trends for Kamala Harris and Donald Trump Across Selected States
harris_data<-harris_data|>mutate(candidate = "Harris")
trump_data<-trump_data |>mutate(candidate= "Trump")
combined_data_total <- bind_rows(harris_data, trump_data)
combined_data_total|>filter(national_poll==1)|>ggplot(aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Country",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-27"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate"))

```
@fig-national-support-trendsillustrates illustrates the polling support trends for Kamala Harris and Donald Trump across the country from August to October. The scatter points represent individual poll results, with blue for Harris and red for Trump, while each dot's size reflects the weight of the poll, accounting for factors like sample size, quality, and recency. Harris’s support trend line generally remains above Trump’s, with an average support level around 49%, compared to Trump’s approximate average of 47% over this period. Toward the end of October, both candidates show a slight increase, with Harris’s support reaching just over 50% and Trump’s approaching 49%. This trend suggests a potential narrowing in support levels as the election approaches, providing a quantitative snapshot of national polling dynamics.

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
# Load necessary libraries
library(boot)    # For bootstrapping
library(caret)   # For cross-validation

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Diagnostics of model using residual vs fitted plot and norm Q-Q plot\-Support for Harris"
#| fig-subcap: ["Residual Plot for Unweighted Model", "Q-Q Plot for Unweighted Model", "Residual Plot for Weighted Model", "Q-Q Plot for Weighted Model"]
#| layout-ncol: 2

@@

Generally, we use Residual vs Fitted plot and Q-Q Plot diagnostic our
model. Residual vs Fitted plot aare Residuals (differences between
observed and predicted values) plotted against fitted values. Ideally,
these residuals should be randomly scattered around the zero line to
indicate that the model does not have systematic errors. The Q-Q plot
for the unweighted model shows how the residuals align with a
theoretical normal distribution. Ideally, residuals should follow a
straight line in this plot if they are normally distributed, which is an
assumption of linear regression.
